---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,        
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.width = 8,
  fig.height = 6,
  dpi = 200,
  fig.align = "center",
  fig.path = "man/figures/README-"
)
knitr::opts_chunk$set()
library(magrittr)
library(queryBuilder)
set.seed(123)
options(tibble.width = Inf)
mtcars <- tibble::as.tibble(mtcars)
options("tibble.print_max" = 5)
options("tibble.print_min" = 5)
pkg_version <- read.dcf("DESCRIPTION", fields = "Version")[1, 1]
```

# queryBuilder

[![version](https://img.shields.io/static/v1.svg?label=github.com&message=v.`r I(pkg_version)`&color=ff69b4)](https://r-world-devs.github.io/queryBuilder/)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

## Overview

`queryBuilder` provides syntax for defining complex filtering expressions in a programmatic way.  
Filtering query, built as a nested list configuration, can be easily stored in other formats like 'YAML' or 'JSON'.
The package also allows to convert such configuration to a valid expression that can be applied with 
popular 'dplyr' package operations.

### Rules

The package allows to construct queries using **rules** (`queryRule`) that are filtering operations performed on a single query.  
A single rule consists of:

- `field` - name of the variable/column to be filtered,
- `operator` - name of the filtering function to apply to the `field`,
- `value` - non-mandatory value precising the filtering by the operator.

As an example:

```{r, eval = FALSE}
queryRule(
  field = "am",
  operator = "equal",
  value = 1
)
```

by the default package configuration, is interpreted as `am == 1` expression.

In order to convert a rule to expression use `queryToExpr` function:

```{r}
my_query <- queryRule(
  field = "am",
  operator = "equal",
  value = 1
)
queryToExpr(my_query)
```

Such expression can be then used by `dplyr::filter`:

```{r}
mtcars %>% dplyr::filter(!!queryToExpr(my_query))
```

To see a full list of supported operators (along with corresponding R functions) check: 
```{r}
listQueryOperators()
```

More detailed description of supported operators can be found at `vignette("operators")`.

You can also define custom operators with `setQueryOperators()`.

### Groups

To build more complex queries `queryBuilder` introduces **groups** (`queryGroup`) that allow to combine multiple rules with the specified **condition** (logical operator).

The below query:

```{r}
my_query <- queryGroup(
  condition = "AND",
  queryRule(
    field = "am",
    operator = "equal",
    value = 1
  ),
  queryRule(
    field = "vs",
    operator = "equal",
    value = 0
  )
)
```

uses `"AND"` condition to combine the two rules which is by default interpreted as `&` logical operator:

```{r}
queryToExpr(my_query)
```

`queryGroup` can also combine other groups which enables to build even more advanced queries:

```{r}
my_query <- queryGroup(
  condition = "AND",
  queryRule("qsec", "greater", 20),
  queryGroup(
    condition = "OR",
    queryRule(
      field = "am",
      operator = "equal",
      value = 1
    ),
    queryRule(
      field = "vs",
      operator = "equal",
      value = 0
    )
  )
)
queryToExpr(my_query)
```

By default the packages supports two conditions `AND` (`&`) and `OR` (`|`) but you can add your custom one with `setQueryConditions()`.

### Relation to [jQuery-QueryBuilder](https://github.com/mistic100/jQuery-QueryBuilder)

The introduced syntax (rules, groups, operators and conditions) is based on query constructing rules as offered by [jQuery-QueryBuilder](https://github.com/mistic100/jQuery-QueryBuilder) JS framework.

The `queryBuilder` package is intended to be used as a backend for the `shinyQueryBuilder` package that will allow users to use `jQuery-QueryBuilder` in Shiny.

## Installation

```{r, eval = FALSE}
# CRAN version
install.packages("queryBuilder")

# Latest development version
remotes::install_github("https://github.com/r-world-devs/queryBuilder")
```

## Acknowledgement

Special thanks to [Kamil Wais](mailto:kamil.wais@gmail.com), [Adam Foryś](mailto:adam.forys@gmail.com), [Maciej Banaś](mailto:banasmaciek@gmail.com),[Karolina Marcinkowska](mailto:karolina_marcinkowska@onet.pl) and [Kamil Koziej](mailto:koziej.k@gmail.com) for the support in the package development and thorough testing of its functionality.

## Getting help

In a case you found any bugs, have feature request or general question please file an issue at the package [Github](https://github.com/r-world-devs/queryBuilder/issues).
You may also contact the package author directly via email at [krystian8207@gmail.com](mailto:krystian8207@gmail.com).
